plyranges: a grammar for transforming genomics data

### 2018-04-20 paper.Rmd
adding data for other examples

bibliography: references.bib
output: 
  rticles::plos_article:
csl: plos.csl
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = "#>", 
                      fig.width = 5, fig.height = 3, fig.align = "center")
options(digits = 2)
library(ggbio)
library(ggplot2)
library(plyranges)

library(AnnotationHub)
q <- AnnotationHub()
# T-cell BW file
bw_file <- q[["AH33458"]] 
```
# Introduction

@@

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(tibble)
feature_tbl <- tribble(
  ~Verb, ~Action, ~Description,
  "`mutate()`", "modifies columns", "Takes a GRanges object and a set of name-value pairs and generates a new GRanges object that with modified or new metadata columns or modified core components (start, end, width, seqnames, strand).",
 "`filter()`", "subset rows", "Takes a GRanges object and a set of logical expressions and restricts the GRanges object to where all logical expressions are true.",
 "`summarise()`", "aggregate one or more columns by a function", "Takes a GRanges object and a set of name-value pairs and aggregates the GRanges according to functions evaluated in the name-value pairs.",
"`group_by()`", "partition a GRanges object by a column", "Create an implicit grouping of GRanges object by one or more columns. This modifies the actions of verbs, so they are performed on each partition created by the grouping.",
"`select()`", "select columns of a GRanges objects", "Drop or reorder metadata columns.",
"`arrange()`", "sort a GRanges object by columns", "Sort a GRanges object by named variables",
"`disjoin_ranges()`", "aggregate one more columns of a GRanges object over the union of end coordinates", "Takes a GRanges object and a set of name-value pairs and expands the GRanges object by taking the union of end coordinates. For each row that was included to form the final union, the aggregation function defined by a name-value pair is called.",
"`reduce_ranges()`", "aggregate one more columns of a GRanges object by merging overlapping and neighbouring ranges.", "Takes a GRanges object and a set of name-value pairs and reduces the GRanges object by merging overlapping and neighouring ranges. For each row that was included to form merged range, the aggregation function defined by a name-value pair is called."
)

>>>>

bibliography: references.bib
output: 
  rticles::plos_article:
csl: plos.csl
---
```{r setup, include = FALSE}
library(tibble)
library(knitr)
opts_chunk$set(message = FALSE, 
               warning = FALSE, 
               comment = "#>", 
               fig.width = 5, 
               fig.height = 3, 
               fig.align = "center")
# for peak figures
library(ggbio)

library(plyranges)

# -- Data from the Human Epigenomics RoadMap consortium
# retrieving a BigWigFile from AnnotationHub
library(AnnotationHub)
q <- AnnotationHub()
# T-cell BW file
bw_file <- q[["AH33458"]]

# a BamFile from the H1 cell line
h1_bam <- "data/GSM433167_BI.H3K27me3.bam"

# Genotype Files accross h1 cell line methylation marks
h1_geno <- "data/GSM1463263_JS-ESCH1.txt.gz"

```
# Introduction

@@

```{r, echo = FALSE, message = FALSE, warning = FALSE}
feature_tbl <- tribble(
  ~Verb, ~Action, ~Description,
  "`mutate()`", "modifies columns", "Takes a GRanges object and a set of name-value pairs and generates a new GRanges object that with modified or new metadata columns or modified core components (start, end, width, seqnames, strand).",
 "`filter()`", "subset rows", "Takes a GRanges object and a set of logical expressions and restricts the GRanges object to where all logical expressions are true.",
 "`summarise()`", "aggregate one or more columns by a function", "Takes a GRanges object and a set of name-value pairs and aggregates the GRanges according to functions evaluated in the name-value pairs.",
"`group_by()`", "partition a GRanges object by a column", "Create an implicit grouping of GRanges object by one or more columns. This modifies the actions of verbs, so they are performed on each partition created by the grouping.",
"`select()`", "select columns of a GRanges objects", "Drop or reorder metadata columns.",
"`arrange()`", "sort a GRanges object by columns", "Sort a GRanges object by named variables",
"`disjoin_ranges()`", "aggregate one more columns of a GRanges object over the union of end coordinates", "Takes a GRanges object and a set of name-value pairs and expands the GRanges object by taking the union of end coordinates. For each row that was included to form the final union, the aggregation function defined by a name-value pair is called.",
"`reduce_ranges()`", "aggregate one more columns of a GRanges object by merging overlapping and neighbouring ranges.", "Takes a GRanges object and a set of name-value pairs and reduces the GRanges object by merging overlapping and neighouring ranges. For each row that was included to form merged range, the aggregation function defined by a name-value pair is called."
)
